---
title: "02_Data_Code"
author: "Kevin Babb"
date: "October 31, 2018"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


The packages "tidyverse" and "broom" should be installed at this stage.
```{r packages, echo=FALSE}
library("tidyverse")
library("broom")
```


And our data loaded as before.
```{r}
raw_stats <- read.csv("~/Documents/Class/CKME-136/Workshop/all_energy_statistics.csv")
test_data <- as_tibble(raw_stats)
test_data <- test_data %>% select(-quantity_footnotes)
```


We check our data:
```{r}
class(test_data)
head(test_data)
```


We filter the categories of interest, beginning with 'Hard coal'.
We drop columns we don't need, group the countries together, and sort the results in ascending order by country followed by year.
Lastly we nest the result by the grouped country.
```{r}
hard_coal <- test_data %>% filter(commodity_transaction == "Hard coal - transformation in electricity, CHP and heat plants") %>% select(-commodity_transaction, -category) %>% group_by(country_or_area) %>% arrange(country_or_area, year) %>% nest()

head(hard_coal)

# Check to see the structure of the 'data' tibble - say Afghanistan
pluck(hard_coal, "data") %>% pluck(1) %>% head()
```



We create new data columns using the 'mutate' and 'map' commands. From the data we extract the following information: 
- initial_year: (first recorded year of transforming this resource), initial_transformation (recorded units of transformation in first recorded year)
- linear model: (derived linear model of transformation units as described by year)
- slope: (slope of linear model: +ve/-ve)
- r_squared: (statistical measure of how close the model data is to the fitted regression line)
```{r}
hard_coal <- test_data %>% filter(commodity_transaction == "Hard coal - transformation in electricity, CHP and heat plants") %>% select(-commodity_transaction, -category) %>% group_by(country_or_area) %>% arrange(country_or_area, year) %>% nest() %>% mutate(initial_year = map_int((map(data, "year")), 1), initial_transformation = map_dbl((map(data, "quantity")), 1), model = map(data, ~lm(quantity ~ year, data =  .)), slope = map_dbl(model, ~pluck(coef(.), "year")), r_squared = map_dbl(model, ~pluck(glance(.), "r.squared")) )

head(hard_coal)
```

We can now begin our analysis on this data. We obtain the a list of the top 20 countries that began with the highest transformtion of coal into electricity.
```{r}
hard_coal %>% arrange(desc(initial_transformation)) %>% head(20)
```


At this point we can generate a chart to see how these countries hard coal transformation into electricity change over time.
```{r}
hard_coal %>% arrange(desc(initial_transformation)) %>% head(20) %>% unnest(data) %>% ggplot(country_or_area, mapping = aes(x = year, y = quantity)) + geom_line(mapping = aes(color = country_or_area))
# We may need to tease this out or do a logarithmic chart to better represent this data.
```
